### 2023-04-21 final/final-daniel-sanchez.Rnw
update final

\title{The influence of entry regulation on formal employment \\[1em] 
\large{ECON 807 Macroeconomic Theory \& Policy Policy Project Draft}\footnote{
Length: 10 pages, not including titlepage, references or appendices. \\
This file has been uploaded with a \texttt{.rar} file including the R project which replicates the environment upon which this paper was written. Also, review the paper's \href{https://github.com/dsanchezp18/econ807-policy-project}{GitHub repository} to quickly glance at the data preparation or the model estimation and data visualization in a \texttt{.Rnw} file (R Sweave/Knitr). The paper's source code is in the \texttt{final} folder in the repository.}}
\author{Daniel Sánchez Pazmiño}
\date{March 2023}

@@

\maketitle
\section{Introduction}

It is well known that informal work -legal but informal economic activities which occur outside the government's regulatory capability \parencite{Sassen.1994}- can hinder economic development through several channels. Informal firms reduce the amount of taxes that can be collected and these businesses tend to remain relativy smaller and less efficient. Informality is also related to bigger gender gaps, higher inequality, lower education and other negative outcomes \parencite{Delechat2020}. Additionally, informal sector workers often lack access to social security and other employment benefits and tend to earn lower wages. The informal economy has been characterized with poorly defined work spaces, unsafe and unsanitary working conditions, long working hours and a lack of access to markets, financial services, training, and technology \parencite{IloND}.

@@

I base my empirical analysis on a theoretical framework informed by the work of \textcite{Branstetter.2014}. In a simple model of entrepreneurship it is predicted that reduced entry costs will increase firm entry and employment, however, all impacts are \textit{at the margin}: new firms will be generally smaller, less productive and less likely to hire large amounts of new workers. While the results from the empirical analysis show a positive effect of the reform on employment, the effect is not robust to different specifications. Further, event study specifications suggest that due to the timing of the policy, the immediate effects of the policy might be due to catch-up effects from the pandemic rather than due to the reform itself. These results suggest that the true causal effect of the policy on employment is too small to identify with this empirical strategy. Some evidence which supports the validity of the model with Ecuadorian data, as the empirical validations in Portugal \parencite{Branstetter.2014} and Mexico \parencite{Kaplan.2011}, also find small positive effects of entry cost deregulation. However, the possibility of a null effect of the reform in Ecuador might be explained by the lack of coordination between entry and labour market regulation.

Through this paper, I contribute to the very reduced stream of research which analyses the impact of entry regulations on employment, considering the existence of informal and formal sectors. Further, the use of administrative instead of survey data enhances the possibilities of extracting insight which is valid at any level of aggregation. No other research links these types of policies to the creation of formal employment, given that these policies are more geared towards fostering entrepreneurship. However, given that \textcite{Branstetter.2014} identifies marginal firms as smaller, which are more likely to be informal \parencite{ElianeElBadaoui.2010}, entry cost regulation may be an important policy tool for tackling informality, especially when labour market reform is politically blocked, as is the case of Ecuador.

>>>>

\title{The influence of entry regulation on formal employment \\[1em] 
\large{ECON 807 Macroeconomic Theory \& Policy Policy Project Final Draft}\footnote{
Length: 12 pages, not including references or appendices. \\
This file has been uploaded with a \texttt{.rar} file including the R project which replicates the environment upon which this paper was written. Also, review the paper's \href{https://github.com/dsanchezp18/econ807-policy-project}{GitHub repository} to quickly glance at the data preparation or the model estimation and data visualization in a \texttt{.Rnw} file (R Sweave/Knitr). The paper's source code is in the \texttt{final} folder in the repository.}}
\author{Daniel Sánchez Pazmiño}
\date{April 2023}

@@

\maketitle

\begin{abstract}

Informal labour markets are very prevalent in Latin America, where often political blockage difficults passing legislation which can address these issues directly. Alternative policies may be taken into account to help tackle the issue of informal work. In this paper, I explore the effect of an entry cost deregulation reform which happened in Ecuador in May 2020. An event-study approach is taken for the empirical analysis with a theoretical framework proposed by \textcite{Branstetter.2014}. While the empirical analysis shows mixed results, there is significant evidence which supports the fit of the theoretical with the data. In this sense, it is possible that a reduction in fixed entry costs for firms can create employment, but such reduction should be large enough to have a statistically significant effect.
\end{abstract}

\section{Introduction}

It is well known that informal work -legal but informal economic activities which occur outside the government's regulatory capability \parencite{Sassen.1994}- can hinder economic development through several channels. Informal firms reduce the amount of taxes that can be collected and these businesses tend to remain relatively smaller and less efficient. Informality is also related to bigger gender gaps, higher inequality, lower education and other negative outcomes \parencite{Delechat2020}. Additionally, informal sector workers often lack access to social security and other employment benefits and tend to earn lower wages. The informal economy has been characterized with poorly defined work spaces, unsafe and unsanitary working conditions, long working hours and a lack of access to markets, financial services, training, and technology \parencite{IloND}.

@@

I base my empirical analysis on a theoretical framework informed by the work of \textcite{Branstetter.2014}. In a simple model of entrepreneurship it is predicted that reduced entry costs will increase firm entry and employment, however, all impacts are \textit{at the margin}: new firms will be generally smaller, less productive and less likely to hire large amounts of new workers. While the results from the empirical analysis show a small positive effect of the reform on employment, the effect is not robust across different specifications. Further, event study specifications suggest that due to the timing of the policy, the immediate effects of the policy might be due to catch-up effects from the pandemic rather than due to the reform itself. These results suggest that the true causal effect of the policy on employment may be too small to identify with this empirical strategy. However, an important and robust finding is that the model seems to support the Ecuadorian data, as it did with the empirical validations in Portugal \parencite{Branstetter.2014} and Mexico \parencite{Kaplan.2011}, who find small positive effects of entry cost deregulation. However, the possibility of a null effect of the reform in Ecuador might be explained by the lack of coordination between entry and labour market regulation, as well as the weaknesses of the present empirical design.

I contribute to the very reduced stream of the literature which analyses the impact of entry regulations on employment, considering the existence of informal and formal sectors. Further, the use of administrative instead of survey data enhances the possibilities of extracting insight which is valid at any level of aggregation. No other research links these types of policies to the creation of formal employment, given that these policies are more geared towards fostering entrepreneurship. However, given that \textcite{Branstetter.2014} identifies marginal firms as smaller, which are more likely to be informal \parencite{ElianeElBadaoui.2010}, entry cost regulation may be an important policy tool for tackling informality, especially when labour market reform is politically blocked, as is the case of Ecuador.
